### 2015-09-04 ms.tex
in introduction, descriptions of previous papers are elaborated

\documentclass[iop,numberedappendix,apj]{emulateapj}

\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{rotating}
\usepackage{natbib}
%\usepackage{lscape}
\usepackage{enumerate}
\usepackage{multirow}
\usepackage{array}
\usepackage{appendix}
\usepackage{comment}
\usepackage{color,xcolor}
\usepackage{url}
\usepackage{hyperref}
\hypersetup{colorlinks,linkcolor={blue!50!black},citecolor={blue!50!black},urlcolor={blue!50!black}}
%\usepackage[dvipdfmx]{color}
%\usepackage[dvipdfmx]{graphicx}

% trivial change as a test

\bibliographystyle{apj}

\def\memoYF#1{\color{red}$[${\bf #1}$]$ \color{black}}

\def\memoDS#1{\color{blue}$[${\bf #1}$]$ \color{black}}

\def\memoNZ#1{\color{green}$[${\bf #1}$]$ \color{black}}

\definecolor{DarkGreen}{rgb}{0.0, 0.6, 0.0}
\def\memoTM#1{\color{DarkGreen}$[${\bf #1}$]$ \color{black}}

\def\plotonesc#1{\centering \leavevmode
\includegraphics[clip=, width=1.70\columnwidth]{#1}}
\def\plotoneh#1{\centering \leavevmode
\includegraphics[clip=, width=.95\columnwidth]{#1}}
\def\plotone#1{\centering \leavevmode
\includegraphics[clip=, width=.85\columnwidth]{#1}}
\def\plotoneShrinkSmall#1{\centering \leavevmode
\includegraphics[clip=, width=.49\columnwidth]{#1}}
\def\plotoneShrinkMed#1{\centering \leavevmode
\includegraphics[clip=, width=.55\columnwidth]{#1}}
\def\plotoneShrinkBig#1{\centering \leavevmode
\includegraphics[clip=, width=.65\columnwidth]{#1}}
\def\plottwo#1#2{\centering \leavevmode
\includegraphics[width=.45\columnwidth]{#1} \hfil
\includegraphics[width=.45\columnwidth]{#2}}
\def\plottwob#1#2{\centering \leavevmode
\includegraphics[width=.49\columnwidth]{#1} \hfil
\includegraphics[width=.49\columnwidth]{#2}}
\def\plottwor#1#2{\centering \leavevmode
\includegraphics[width=.55\columnwidth,angle=90]{#1} \hfil
\includegraphics[width=.55\columnwidth,angle=90]{#2}}
\def\plotthree#1#2#3{\centering \leavevmode
\includegraphics[width=.3\columnwidth]{#1} \hfil
\includegraphics[width=.3\columnwidth]{#2} \hfil
\includegraphics[width=.3\columnwidth]{#3}}

\newcommand{\cN}[1]{\mathcal{N}}
\newcommand{\pn}[1]{\mbox{$(#1)$}}
\newcommand{\spa}{\mbox{ }}
\def\gsim{\;\rlap{\lower 2.5pt
 \hbox{$\sim$}}\raise 1.5pt\hbox{$>$}\;}
\def\lsim{\;\rlap{\lower 2.5pt
   \hbox{$\sim$}}\raise 1.5pt\hbox{$<$}\;}

\newcommand{\Tcmb}{\mbox{$T_{\mbox{\tiny CMB}}$}}
\newcommand{\Tsky}{\mbox{$T_{\mbox{\tiny sky}}$}}
\newcommand{\Tsys}{\mbox{$T_{\mbox{\tiny sys}}$}}
\newcommand{\Trx}{\mbox{$T_{\mbox{\tiny rx}}$}}
\newcommand{\sigRMS}{\mbox{$\sigma_{\mbox{\tiny RMS}}$}}

% set formatting properties
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{8.8in}
\setlength{\hoffset}{0.0in}
\setlength{\voffset}{-0.4in}
%\setlength{\voffset}{0.3in}
\parindent 0.2in
\parskip 0.1in

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% THE DOCUMENT BEGINS HERE                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\slugcomment{Submitted to ApJ, 20 October 2011}

\begin{document}

%%% Begin front material
%\twocolumn[%%% Begin front material

\title{Red-Giant Hot Jupiters: Brilliant Radio Beacons?}

\author{
%
Yuka Fujii\altaffilmark{0} \\
%
David S. Spiegel\altaffilmark{1,2} \\
%
{\bf and some order:} \\
%
Jason Nordhaus\altaffilmark{3,4} \\
%
Nikku Madhusudhan\altaffilmark{5} \\
%
Mehrdad Mirbabayi\altaffilmark{1} \\
%
Aaron Parsons\altaffilmark{6} \\
%
Tony Mroczkowski\altaffilmark{7,8} \\
%
Neil T. Zimmerman\altaffilmark{9}
}

\affil{$^0$Earth-Life Science Institute, Tokyo Institute of Technology, 
  Tokyo, JAPAN, 152-8550}

\affil{$^1$Astrophysics Department, Institute for Advanced Study,
  Princeton, NJ 08540}

\affil{$^2$Research \& Development, Sum Labs,
  New York, NY  10001}

\affil{$^3$Department of Science and Mathematics, National Technical Institute for the Deaf, Rochester Institute of Technology, Rochester, NY 14623}

\affil{$^4$Center for Computational Relativity and Gravitation, Department of Mathematics, Rochester Institute of Technology, Rochester, NY 14623}

\affil{$^5$Astronomy Department, University of Cambridge, UK}

\affil{$^6$Astronomy Department, University of California Berkeley}

\affil{$^7$National Research Council Fellow}

\affil{$^8$Naval Research Laboratory, 4555 Overlook Ave SW, Washington, DC 20375}

\affil{$^9$Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218}

\vspace{0.5\baselineskip}

\email{
yuka.fujii@elsi.jp
}

\begin{abstract}
%%%
When planet-hosting stars evolve off the main sequence and go through the red-giant branch (RGB), the stars become orders-of-magnitudes more luminous and at the same time lose mass at much higher rates than their main-sequence counterparts.
Accordingly, planetary companions around them at a few AU orbits, if exist, are heated up to the level of canonical hot Jupiters and also subjected to the massive stellar wind.
Given that magnetized planets emit radio waves in the interaction with stellar winds, such ``Red-Giant Hot Jupiters'' (RGHJs) may be candidates of bright radio emitters.
We estimate the spectral auroral radio intensity of RGHJs employing the ``radiometric Bode's law'' and proposed scaling of planetary magnetic fields.
We find that RGHJs might be as bright as canonical hot Jupiters as predicted so far, and thus radio emission from a RGHJ would be visible from a hundred parsec distances with current/future detectors such as LOFAR-HBA and SKA.

%Red-giant hot Jupiters (RGHJs) are jovian planets orbiting red-giant-branch (RGB) or asymptotic-giant-branch (AGB) stars. 
%Red-giant hot Jupiters (RGHJs) are jovian planets orbiting red-giant-branch (RGB) or asymptotic-giant-branch (AGB) stars.
%Post-main-sequence stars lose mass at much higher rates than main-sequence stars.
%A jovian planet passing through the dense winds of its RGB or AGB host can capture stellar wind in its magnetosphere, leading to radio emission from electrons spiraling in the magnetic field.
%The cyclotron frequency of electrons from the stellar wind accreting onto the planet scales as 100~MHz~$(B/30 {\rm~Gauss})$.
%The cyclotron frequency of electrons from the stellar wind accreting onto the planet scales as 100~MHz~$(B/30 {\rm~Gauss})$.
%Counterintuitively, a plausible toy model suggests that a stronger planetary magnetic field leads to lower spectral flux density, and might therefore make objects more challenging to observe.
%Still, for a range of planetary and stellar parameters, a RGHJ might generate a radio luminosity that would be visible from kiloparsec distances.
\end{abstract}

\keywords{planets and satellites: Jupiter --- Sun: evolution ---
  planetary systems --- stars: evolution ---
  stars: AGB and post-AGB}
%]%%% End front material

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Planets with strong magnetic fields may generate radio or X-ray emission when interacting with energetic charged particles. 
It is well known that Jupiter emits radio waves from its auroral region due to the cyclotron-maser instability \citep{zarka1998}.  % due to the acceleration of the plasmas. \memoYF{?}.
Potentially, exoplanets can also generate radio emission through similar mechanisms, depending on their intrinsic magnetic fields and the density of surrounding plasmas, e.g. stellar wind particles and particles from Io-like moons. 
Observations of radio emission from Solar System planets imply an empirical relation that the radio intensity is proportional to the input solar wind flux going into each planetary magnetosphere, which is commonly referred to as the ``radiometric Bode's law'' \citep{desch+kaiser1984}. 
Extrapolating this scaling to exoplanets, radio emission from exoplanetary systems have been examined. 
\citet{farrell1999,zarka2001,lazio2004} pointed out that a few of the known exoplanetary systems may have radio flux of $\sim $1 mJy level, due to the small orbital distance of the planet and/or the larger planetary mass. 
%\citet{griesmeier2004} modeled in detail the evolution of atmosphere and magnetic field of the tidally-locked close-in stars and found that 
\citet{stevens2005} gave improved estimates of the stellar mass-loss rate based on X-ray flux and re-evaluated the radio flux of known exoplanetary systems. 
\citet{griesmeier2005} took account of the high stellar activity in the early stage of the planetary system and proposed that the young system would be good candidates to search for radio emission. 
\citet{griesmeier2007a, griesmeier2007b} discussed the effects of the detailed model of stellar wind property in the proximity of the stars. They considered not only the kinetic energy of the stellar wind but also the magnetic energy of the stellar wind and coronal mass ejection (CME). 
Note that in these paperes, the scaling relation of planetary magnetic fields  differ from paper to paper.    
In \citet{reiners2010}, the authors adopted new scaling relation of planetary magnetic field based on \citet{christensen2009}.
\citet{jardine2008} modeled the reconnection between the magnetic fields of close-in planets and of the star, which would accelerate electrons and produce radio emission.  
Although observational searches for these radio signatures are underway, no clear detection has been claimed \citep{bastian2000,george2007,stroe2012,hallinan2013,murphy2015}, 
though there are some promising initial results \citep{lecavelier_et_al2013,sirothia2014}.

When stars less than $\sim$8~$M_\sun$ evolve off the main sequence,
they evolve through the red-giant branch (RGB) and the
asymptotic-giant branch (AGB) phases where their radii and luminosities increase by orders of magnitude. 
Jovian planets in orbit around such stars can be transiently heated to hot-Jupiter temperatures ($\sim$1000~K or more) at distances out to tens of AU, depending on the star's mass; such planets are termed ``Red Giant Hot Jupiters'' (RGHJs) but refer to planets orbiting RGB and AGB stars \citep{spiegel+madhusudhan2012}. 
Such planets are also subject to interactions with a massive (but slow) stellar wind, as the mass-loss rate of evolved stars is significant, ranging from $\sim  10^{-8} M_\odot$/yr to $\sim  10^{-5} M_\odot$/yr with the highest values for AGB stars \citep[e.g.,][]{reimers1975, schild1989, vassiliadis1993, schoier2001, vanloon2005}, thus higher than that of the Sun, $\sim 10^{-14}M_\odot$/yr. 
On the assumption that the radio emission is correlated with the stellar wind, planetary companions around evolved stars could also generate bright radio emission.

Based on this speculation, \citet{ignace2010} examined radio emission from known substellar-mass companions around cool evolved stars.
They found that the low ionization fraction of the stellar wind of evolved stars suppresses their radio emission and leads to weak radio emission, probably below the detection limit.

In this paper, we argue that the accretion of the massive stellar wind onto the planet would emit strong X-rays, thereby ionizing the hydrogen atoms around the planet.
The ionized stellar wind would then interact with the planetary magnetic field in the same way as Solar System planets do.
Thus, by extrapolating the radiometric Bode's law, we provide more optimistic estimates for highly evolved stars compared to the previous result.
Given that the survey of exoplanets around highly evolved stars is not complete, we do not constrain ourselves to the systems known to have sub-stellar companions.
Instead, we investigate the parameter space where we can search for radio emission and examine the possibility to detect them with current and future instruments.

In \S2, we describe our assumptions for the stellar wind, planetary magnetosphere, frequency of radio emission, and scaling of planetary radio emission.
\S3 presents an estimate of the spectral radio intensity of RGHJs and compares the predictions with what might be expected from canonical hot Jupiters as well as those from Jupiter-twins.
\S4 gives predictions for the signals from evolved systems within 400 pc, taking account of the intrinsic radio emission from the host evolved stars. Considering the sensitivity of the current and future radio instruments, we argue that SKA, and perhaps LOFAR-HBA as well, would be capable of detecting these signals within a reasonable integration time. 
%In \S5, we discuss the usefulness of radio emission in exploring Jupiter-like planets around evolved stars. 
In \S5, we mention the effects of canceling magnetic field which results from the plasmas spiraling in the planetary magnetosphere.  
Finally, \S6 concludes the paper with a brief summary.

%Roughly 20\% of the more than 700 \memoYF{check!} currently known exoplanets around main-sequence stars \footnote{See online catalogs such as http://www.openexoplanetcatalogue.com/ \citep{rein2012}, http://exoplanet.eu \citep{schneider_et_al2011}, or http://exoplanets.org \citep{wright_et_al2011} for up-to-date lists.} have masses greater than half of Jupiter's, orbital radii greater than 1~AU, and will become hot Jupiters (i.e., for the present purposes, this means they will receive at least as much irradiation as the currently known hot Jupiters/Neptunes \citep{spiegel+madhusudhan2012}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Model}
\label{s:assumptions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\memoYF{In the end the number of significant decimal digits is set to 1 or 2.}

Before estimating the frequency of planetary radio emission and the spectral radio intensity, we specify the property of the planetary magnetic field in \S\ref{ss:magneticfield}, and that of the stellar wind of the evolved stars in \S\ref{ss:stellarwind}.  
Then, we discuss the framework to estimate the frequency and intensity of planetary radio emission in \S\ref{ss:model_frequency} and \S\ref{ss:model_intensity}, respectively.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Assumptions for Planetary Magnetic Field}
\label{ss:magneticfield}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We estimate the planetary magnetic field of general planets  
simply by scaling the Jovian magnetosphere at the surface $B_J \sim 10$ G according to the planetary mass and age, based on the predicted scaling relation described below. 
% we adopt a simple scaling relation between the magnetic field and macroscopic planetary properties.

So far, several scaling relations for the planetary magnetic field strength  
have been proposed \citep[e.g.][]{russel1978,busse1976,stevenson1979,mizutani1992,sano1993,starchenko2002,christensen2006,christensen_et_al2009} %\memoYF{check!}
%(see \citet{griesmeier2004} or \citet{christensen2010} for the summary), 
which are summarized and compared with numerical simulations in \citet{christensen2010}. 
We employ the scaling law proposed in \citet{christensen_et_al2009}, which has already been used in \citet{reiners2010} to explore the evolution of planetary magnetic fields.
This scaling law is based on the assumption that the ohmic dissipation energy is a fraction of available convected energy and was found to be in a good agreement with the numerical experiments over a wide parameter space and with known objects from the Earth to stars: 
%\memoYF{describe in more detail}:
%%%%%%%%%% 
%\begin{equation}
%B^2 \propto \rho_c r_c^{4/3} q_c^{2/3}, \label{eq:Bscaling} 
%\end{equation}
\begin{equation}
B_{\rm dynamo} ^2 \propto f_{\rm ohm}\;\rho_{\rm dynamo}  ^{1/3} \;  (Fq_o)^{2/3} \label{eq:Bscaling} 
\end{equation}
%%%%%%%%%%
%where $\rho _c$ and $r_c$ are the density and the radius of the dynamo region, and $q_c$ is the internal convected energy flux in the core region. 
where $B_{\rm dynamo}$ is the mean magnetic field in the dynamo region, $f_{\rm ohm}$ is the ratio of ohmic dissipation to the total dissipation, $\rho_{\rm dynamo} $ is the mean density in the dynamo region, $F$ is an efficiency factor of order unity, and $q_o$ is the convected flux at the outer boundary of the dynamo region \citep[see][for the comprehensive description]{christensen_et_al2009}. 
Here, $f_{\rm ohm}$ and $F$ are assumed to be constant for the bodies considered in this paper. 
Dipole magnetic field strength at the planetary surface, denoted by $B$, is then scaled by
%%%%%%%%%%
\begin{equation}
B \propto B_{\rm dynamo} \left( \frac{r_{\rm dynamo}}{R_p} \right)^3.  \label{eq:Bscaling2}
\end{equation}
%%%%%%%%%%
where $r_{\rm dynamo}$ is the radius of the outer boundary of the dynamo region.

Note that the scaling law, Equation (\ref{eq:Bscaling}), would be reasonable only for rapidly rotating objects.
Indeed, unlike canonical hot Jupiters, RGHJs are not tidal locked to their host stars, so we do not have to assume slow spin rotation rates \citep{spiegel+madhusudhan2012}.
Therefore, we implicitly assume that the RGHJs we consider in this paper are rapidly rotating so that they generate planetary magnetic fields through the same mechanism as Jupiter.

In order to evaluate $\rho _{\rm dynamo}$ and $q_o$, we need models of the internal planetary structure. 
We consider Jupiter-like gaseous planets and assume that the planetary radius is constant at $R_p = R_{p,J}$, as numerical calculations show that the radii of gaseous planets over the range of $0.1 M_{p, J} < M_p < 10M_{p, J}$ (with core mass less than 10\%) are converged to $0.8 R_{p, J} < R_p < 1.2R_{p, J}$ within 1 Gyr \citep{fortney2007}. 
For the density profile, we assume a polytrope gas sphere with index $n=1$, which results in:
%%%%%%%%%% 
\begin{equation}
\rho [r] = \left( \frac{\pi M_p}{4 R_p^3} \right) \frac{\sin \left[ \pi \frac{r}{R_p} \right]}{\left( \pi \frac{r}{R_p} \right)}. \label{eq:rho_r}
\end{equation}
%%%%%%%%%%
We determine the radius of outer boundary of the dynamo region, $r_{\rm dynamo}$, by assuming that the hydrogen becomes metallic when $\rho (r)$ exceeds the critical density $\rho_{\rm crit}=0.7\,\mbox{g/cm}^3$ \citep{exoplanets2006, griesmeier2007b}.
The density of the metallic core, $\rho _{\rm dynamo}$ is obtained by averaging the density in the core. 
In the case of Jupiter, $r_{\rm dynamo, J} = 0.85 R_{\rm J}$ and $\rho_{\rm dynamo, J} = 1.899~{\rm g/cm}^3$.

The scaling of convected heat flux at the outer boundary, $q_o$, is obtained by dividing the age-dependent net-planetary luminosity by the surface area of the core region, i.e., $4\pi r_{\rm dynamo}^2$. 
The time-dependence luminosity is taken from \citet{burrows_et_al2001} \citep[see also][]{marley2007}. 
Ignoring the relatively weak dependence on average atmospheric Rosseland mean opacity leads to:
%%%%%%%%%%
\begin{equation}
L \sim 1.84\times10^{-10} L_\odot \left( \frac{t}{4.5 \rm~Gyr} \right)^{-1.3} \left( \frac{M_p}{M_J} \right)^{2.64} \, 	
\label{eq:burrowsLum}
\end{equation}
%%%%%%%%%%
Therefore, we have
%%%%%%%%%%
\begin{equation}
q_o \sim q_{o, J} \left( \frac{t}{4.5 \rm~Gyr} \right)^{-1.3} \left( \frac{M_p}{M_J} \right)^{2.64} \left( \frac{r_c}{r_{c,\,J}} \right)^{-2}\, .
\label{eq:burrowsHeatFlux}
\end{equation}
%%%%%%%%%%

Substituting Equation~\ref{eq:burrowsHeatFlux} into Equation~\ref{eq:Bscaling2} and given that $r_{\rm dynamo} $ does not change significantly, the magnetic field is roughly, we find that:
%%%%%%%%%% 
\begin{equation}
%B   &\sim &  10~{\rm G} \left( \frac{t}{4.5~\rm{Gyr}} \right)^{-0.43} \left( \frac{M_p }{M_{p,J}} \right)^{1.38} \label{eq:scalingB} \\
B   \sim   10~{\rm G} \left( \frac{M_p }{M_{p,J}} \right)^{1.04} \left( \frac{t}{4.5~\rm{Gyr}} \right)^{-0.43} \label{eq:scalingB}
\end{equation}
%%%%%%%%%%
under this model.

%We note that the scaling relation based on \citet{christensen_et_al2009}, or equation (\ref{eq:Bscaling}), has also been adopted in \citet{reiners2010}. While the detailed treatment of $\rho _c$ and $q_o$ are different, two estimates of planetary magnetic field strength are fairly close, as shown in \S3 below.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Assumptions for Stellar Wind}
\label{ss:stellarwind}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Another ingredient needed for modeling the radio emission from the interaction between the planetary magnetic field and the stellar wind is the density and velocity of the stellar wind. 
The number density of electrons in the stellar wind, $n$, can be expressed as
%%%%%%%%%% 
\begin{equation}
n = \frac{\dot M_\star}{4\pi a^2 m_p v}
\label{eq:n}
\end{equation}
%%%%%%%%%% 
where $\dot M_\star$ is the stellar mass loss rate, $a$ is the orbital distance from the star and $m_p$ is proton mass, and $v$ is the velocity of the stellar wind.

While the solar mass loss rate, $\dot M_\odot$, is $\dot M_\odot \sim 2\times 10^{-14} M_{\odot}$/yr \citep[e.g.,][]{hundhausen1997}, the mass loss rate of red giants is typically $\dot M_\star \sim 10^{-8}-10^{-7} M_{\odot}$/yr \citep{reimers1975}, and the rate can be as high as $10^{-5} M_\odot{\rm /yr}$ during the AGB phase \citep{schild1989, vassiliadis1993, schoier2001, vanloon2005}.
Therefore, we have
%%%%%%%%%% 
\begin{equation}
\frac{\dot M_\star}{\dot M_{\odot}} \sim 10^6 - 10^9 \, . \label{eq:scale_Mdot}
\end{equation}
%%%%%%%%%%

On the other hand, the stellar wind velocity becomes smaller because of the small escape velocity (due to the expanded stellar radius).
%\footnote{Escape velocity is $\sqrt{2GM/r}$.}
Assuming that the stellar wind scales as the escape velocity, i.e., $\propto \sqrt{2GM_{\star}/R_{\star}}$ ($G$ is the gravitational constant, $M_{\star}$ is the stellar mass, and $R_{\star}$ is the stellar radius), a Solar-mass red giant with radius $R_{\star}=100R_{\odot}$ encounters 10 times as slow stellar wind as that at the Sun, $v_{\odot}$.
Therefore, 
%%%%%%%%%% 
\begin{equation}
\frac{v}{v_{\odot}} \sim 10^{-1} \, . \label{eq:scale_v}
\end{equation}
%%%%%%%%%% 
%where $v_{\odot} \sim 4.0 \times 10^2 $ km/sec.

Based on equations (\ref{eq:scale_Mdot}) and (\ref{eq:scale_v}), the number density of the stellar wind (equation (\ref{eq:n})) are normalized as follows:
%%%%%%%%%% 
\begin{eqnarray}
n &=& 1.80 \times 10^6 \; {\rm cm^{-3}} \\
&&\times \left( \frac{\dot M_\star}{10^{-8} M_{\odot }{\rm /yr}} \right) \left( \frac{v}{30\; {\rm km}} \right)^{-1} \left( \frac{a}{5 \; {\rm AU}} \right)^{-2} \label{eq:n_normalized}
\end{eqnarray}
%%%%%%%%%%

\memoYF{check from here}
As discussed in \citet{ignace2010}, one potential problem of the stellar wind of evolved stars which is unfavorable for radio emission is that their ionization fraction is usually as low as $10^{-3}$ \memoYF{cite something}, which would lead to inefficient interaction with planetary magnetosphere and hence low input energy for radio emission. 
However, at this stage the velocity of stellar wind becomes slower than the escape velocity of planetary companion and hence the stellar wind particles are expected to accrete onto the planets. 
Based on \citet{spiegel+madhusudhan2012}, the accretion luminosity $L_{\rm acc}$ and the temperature $T_{\rm acc}$ are: 
%%%%%%%%%
\begin{eqnarray}
L_{\rm acc} \sim &&  10^{29} \; {\rm erg\;s}^{-1} \\
&& \times \left( \frac{\dot M_*}{10^{-7} M_{\odot }\;{\rm yr}} \right) \left( \frac{M_p}{10 M_J} \right)^3 \left( \frac{M_*}{M_{\odot }} \right) \\
%
T_{\rm acc} \sim && 2 \times 10^6 \left( \frac{M_p}{10 M_J} \right)
\end{eqnarray}
%%%%%%%%%
The accretion disks around planets therefore emit X-ray radiation. 
We estimate the scale of the ionized region by computing Stromgren radius of the accretion disk: 
%%%%%%%%%
\begin{equation}
R_{\rm stromgren} = \frac{3}{4\pi} \left( \frac{S}{n^2 \beta } \right)^{1/3}
\end{equation}
%%%%%%%%%
where $S$ is the source rate of ionizing photons, $n$ is the density of hydrogen, and $\beta $ is the coefficient for recombination \memoYF{need to check}:
% http://www.tapir.caltech.edu/~chirata/ay102/Photoionized.pdf
% T \propto 4.3e-13 ( T /10^4 K )^-0.7
%%%%%%%%%
\begin{eqnarray}
\beta &=& 4.3? \times 10^{-13} \; \left( \frac{T}{10^4 \;{\rm K}} \right) ^{-3/4} \;\; {\rm cm^3 \; sec^{-1}} \\
&\sim& 5.28 \times 10^{-14} \; \left( \frac{M_p}{M_J} \right)^{-3/4} \;\; {\rm cm^3 \; sec^{-1}} 
\end{eqnarray}
%%%%%%%%%
The source rate is obtained by counting the number of photons whose energy is higher than 13.6~eV, the energy needed to ionize hydrogen atoms.
As a result, \memoYF{need to check}
%%%%%%%%%
\begin{eqnarray}
S_* = 
  \left\{
    \begin{array}{ll}
      1.18 \times 10^{42} \;  {\rm sec^{-1}} & \;\;\;(M_p = M_J) \\
      2.66 \times 10^{44} \; {\rm sec^{-1}} & \;\;\;(M_p = 10 M_J)
    \end{array}
  \right.
\end{eqnarray}
%%%%%%%%%
and the number density of hydrogen atoms are given by equation (\ref{eq:n_normalized}). 
These values give \memoYF{need to check}
%%%%%%%%%
\begin{eqnarray}
R_{\rm stromgren} = 
  \left\{
    \begin{array}{ll}
      3 \;{\rm AU} & \;\;\;(M_p = M_J) \\
      30 \;{\rm AU} & \;\;\;(M_p = 10 M_J)
    \end{array}
  \right.
\end{eqnarray}
%%%%%%%%%

We therefore expect that the X-ray emission from the accretion disk is enough to ionize the stellar wind.
This implies that the stellar wind particles represented by (equation~\ref{eq:n}) are fully ionized plasmas like Solar wind, and that the stellar wind can interact with the planetary magnetic fields in a rather similar way to the case of Solar System planets.

%\citet{drake1983}
%\citet{}
%\color{black}

%The temperature of stellar wind of RGs is expected to be 2 orders of magnitude lower than their main sequence counterparts, mainly because the sound speed of hot corona exceeds the escape speed, i.e., hot corona cannot be confined in the stellar atmosphere \citep{suzuki2008}. 
%\begin{equation}
%\frac{T}{T_{\odot}} \sim 10^{-2}
%\end{equation}
%where $T_{\odot } \sim 10^6$ K.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Frequency of Radio Emission}
\label{ss:model_frequency}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Planetary auroral radio wave are emitted at the local cyclotron frequency along the motion of the plasmas.
The upper limit is around the cyclotron frequency of the planetary surface magnetic field, $f_{\rm cyc}$: 
%%%%%%%%%% 
\begin{equation}
f_{\rm cyc} = \frac{eB}{2\pi m_e} \approx 27.9 {\rm~MHz} \left( \frac{B}{10 \rm~G} \right) \label{eq:fcyc}
\end{equation}
%%%%%%%%%%
where $e$ and $m_e$ are the charge and mass of the electron, respectively.

Radio emission from the planet is observable from ground only when the maximum frequency is larger than the plasma frequency of Earth's ionosphere $f_{\rm plasma}^\oplus$ and the maximum plasma frequency along the line of sight $f_{\rm plasma}^{\rm los}$.
The plasma frequency may be expressed as
\begin{eqnarray}
f_{\rm plasma} & = & \sqrt{\frac{n_e e^2}{\pi m_e}} \\
 & = & 8979 {\rm~Hz} \times \left( \frac{n_e}{\rm cm^{-3}} \right)^{1/2} \, .
\label{eq:fplasma}
\end{eqnarray}
In the Earth's ionosphere, the electron number density is less than $10^6$~cm$^{-3}$, which implies that $f_{\rm plasma}^\oplus \lsim 10$~MHz.
At the favorable viewing condition, when the planet is on the near side of its star to the Earth, the maximum plasma frequency along the line of sight corresponds to that in the vicinity of the planet.
Therefore, combining Eq.~(\ref{eq:n}) and Eq.~(\ref{eq:fplasma}), we have
%%%
\begin{eqnarray}
\label{eq:fplasmalos} f_{\rm plasma}^{\rm los} & \sim & 8.98 {\rm~kHz} \times \left( \frac{\dot M_\star}{4 \pi a^2 m_p v} \times 1\rm~cm^3 \right)^{1/2} \\
\label{eq:fplasmalos_scaled} & = & 14.7{\rm~MHz} \times \left( \frac{a}{5~\mbox{AU}}\right)^{-1} \left(\frac{v}{0.1 v_{\odot}}  \right)^{-1/2} \notag \\
 & & \times \left( \frac{\dot M_\star}{10^6 \dot M_{\odot}}\right)^{1/2} \, .
\end{eqnarray}
%%%
%where in Eq.~(\ref{eq:fplasmalos_scaled}) we have taken the solar mass-loss rate $\dot M_\odot$ to be $2.0 \times 10^{-14}~M_\odot$/yr and the typical solar wind velocity $v_{\odot }$ to be 400~km/sec.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Flux of Radio Emission}
\label{ss:model_intensity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The auroral radio spectral flux of exoplanets observed at the Earth, $F_{\nu}$, can be expressed by:
%%%
\begin{equation}
F_{\nu} = \frac{P_{\rm radio}}{\Omega l^2 \Delta \nu}
\label{eq:Fnu}
\end{equation}
%%%
where $P_{\rm radio}$ is the energy that is deposited as radio emission of considered frequency range, $\Omega$ is the solid angle of the emission, $l$ is the distance between the target and the Earth, and $\Delta \nu$ is frequency bandwidth.

We estimate the radio emission of exoplanets, $P_{\rm radio}$, simply by scaling the Jovian auroral radio emission, $P_{\rm radio,\,J}$, with the input energy from stellar wind, in the same manner as \citet{griesmeier2005,griesmeier2007a,griesmeier2007b}.
The scaling is based on the empirical/apparent good correlation between the radio emission intensity of Solar System planets and the input kinetic energy, $P_{\rm inp,\,k}$, or the magnetic energy of the solar wind, $P_{\rm inp,\,m}$, \citep[``radio Bode's law''; ][]{desch+kaiser1984}, i.e.,
%Namely, although the detail of the mechanism to generate planetary radio emission is not fully understood so far \memoYF{cite something?}, 
%observations indicate that the total radio emission, $P_{\rm radio}$ is proportional to the input kinetic energy, $P_{\rm k,\,inp}$, or the input magnetic energy, $P_{\rm m,\,inp}$ \citep[``radio Bode's law''; ][]{desch+kaiser1984}:
%%%
\begin{eqnarray}
P_{\rm radio} &\propto & P_{\rm inp} \\
P_{\rm inp,\,k} &=& n v^3 r_s ^2 \label{eq:Pinp_kin}, \\
P_{\rm inp,\,m} &=& v B_{\bot }^2 r_s ^2 \label{eq:Pinp_mag} \, ,
\end{eqnarray}
%%%
where $ B_{\bot }$ is the interstellar magnetic field perpendicular to the stellar wind flow, and $r_s$ is the radius of the magnetic stand-off point. 
%In the case of the Solar wind, the kinetic energy is dominant input energy over the magnetic one (the former is $\sim 400$ times larger than the latter).  
It is not clear from the observations of the Solar System planets which is the fundamental one, the correlation to the input kinetic energy, that to the magnetic energy, or both. 
In this paper, we assume that the radio emission is scaled with the input kinetic energy, i.e., $P_{\rm radio} \propto P_{\rm k,\,inp}$ and consider the possible effects of massive stellar wind.

On the other hand, in the case that the correlation with input magnetic energy is fundamental, it is difficult to estimate the planetary radio emission at this point because magnetic field of evolved stars are not well constrained for highly evolved stars; most of observations only set an upper limit \citep[e.g.,][]{konstantinova2010,petit2013,tsvetkova2013,konstantinova2013,auriere2015}. 
In the case of M-type giant EK boo, surface magnetic field $\sim 0.1-10 $G has been measured; In this particular case, given the large stellar radius $R_\star \sim 210 R_{\odot }$ the magnetic moment may be about $10^3$ times larger than the Sun and thus may also increase the planetary radio emission.
However, because of the lack of universal understanding of the stellar magnetic field, we leave the magnetic model of radio emission for evolved stars for future work.

In reality, the total power of Jovian auroral emission varies greatly over time: $1.3\times 10^{10}$ W for the average, $3.2\times 10^{10}$ W for the average of highly active time, and $4.5 \times 10^{11}$ W for the peak activity \citep{zarka_et_al2004}. 
Here we employ $P_{\rm radio,\,J}=2.1\times 10^{11}\mbox{~W}$ as a canonical value, following \citet{griesmeier2005,griesmeier2007b}.
\memoYF{reconsidering...}

The radius of the magnetic stand-off point, $r_s$, is obtained based on the balance between the stellar wind pressure and the planetary magnetic pressure: 
%%%
\begin{equation}
m_p n v ^2 \sim \frac{B^2}{8\pi}\left( \frac{r_s}{R_p} \right)^{-6}  \label{eq:stand-off}
\end{equation}
%%%
The radius obtained for Jupiter from this equality is about a half of actual radius of magnetosphere \citep[][]{griesmeier2005}. 
By scaling this radius with equations (\ref{eq:n}) and (\ref{eq:stand-off}), the radius of magnetosphere of planets would be estimated as:
%If the radius of magnetosphere is estimated by scaling that of Jupiter according to equations (\ref{eq:n}) and (\ref{eq:stand-off}), 
%We estimate $r_s$ simply by scaling Jovian magnetosphere \citep[$r_{s,\,J} \sim 84 R_J$;][]{joy2002} according to equation (\ref{eq:stand-off}), i.e., 
%%%
\begin{eqnarray}
\nonumber r_s 
%&\sim & R_p \left( \frac{B^2}{8\pi m_{\rm p}nv^2} \right)^{1/6}  \\
&=& r_{s,\,J} \left( \frac{B}{10~\mbox{G}} \right)^{1/3} \left( \frac{a}{5.2\mbox{AU}} \right)^{1/3}  \\
&& \times \left( \frac{\dot M_\star}{\dot M_\odot} \right)^{-1/6}  \left( \frac{\dot v}{v_{\odot }} \right)^{-1/6} \\
\nonumber &=& 11.3 \, R_J \left( \frac{B}{10~\mbox{G}} \right)^{1/3}  \left( \frac{a}{5\mbox{AU}} \right)^{1/3} \\
&& \times \left( \frac{\dot M_\star}{10^6 \dot M_\odot} \right)^{-1/6}  \left( \frac{\dot v}{10^{-1}v_{\odot }} \right)^{-1/6}
 \label{eq:stand-off-radius}
\end{eqnarray}
%%%

%Thus, the radio emission, $P_{\rm radio}$, or equivalently the input kinetic energy, $P_{\rm k,\,inp}$, are estimated with $n$, $v$, and $\mathcal{M}$, which should be specified by assuming the properties of the stellar wind and the planetary magnetic field. 
%In the following, we describe our assumptions for these properties.

We assume that the solid angle of the emission is same as the Jupiter's one for all of the exoplanets.
In reality, the solid angles of auroral radio emission from  Jupiter, Saturn, and Earth are $\sim 1.6$, $\sim $6.3, and $\sim $3.5, respectively \citep{desch+kaiser1984}, which are on the same order and will not significantly affect our order-of-magnitude estimate of radio emission. 
The bandwidth, $\Delta \nu$, is assumed to be proportional to the representative frequency of the emission, which is the cyclotron frequency, following \citet{griesmeier2007b}.

%We estimate the radio brightness of RGHJs by simply scaling \memoDS{stray sentence fragment?}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\label{s:result}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Estimate of Planetary Magnetic Field and Maximum Frequency}
\label{ss:Bplanet}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%s

Figure \ref{fig:planetaryB_pre} shows the computed radius ($r_c$) and average density ($\rho_c$) of the dynamo region, and Figure \ref{fig:planetaryB} displays the estimates of the strength of planetary magnetic field $B$ based on the scaling relation in \S\ref{ss:magneticfield} and the corresponding cyclotron frequency ($f_{\rm cyc}$), as functions of planetary mass and the age. 
%Given that $\rho _c \propto M_p$  at large $M_p$ and $q_c \propto t^{-1.3} M_p^{2.64}/r_c^2$, the magnetic field is roughly:
Reasonably, the resultant values are in a good agreement with \citet{reiners2010}, which adopted the same scaling law for planetary magnetic field; we show this figure just for self-completeness.

%%%%%%%%%% 
%\begin{eqnarray}
%B   &\sim &  10~{\rm G} \left( \frac{t}{4.5~\rm{Gyr}} \right)^{-0.43} \left( \frac{M_p }{M_{p,J}} \right)^{1.38} \label{eq:scalingB} \\
%B   &\sim &  10~{\rm G} \left( \frac{t}{4.5~\rm{Gyr}} \right)^{-0.43} \left( \frac{M_p }{M_{p,J}} \right)^{1.04} \label{eq:scalingB} \\
%f_{\rm cyc} &\sim &  28~{\rm MHz} \left( \frac{t}{4.5~\rm{Gyr}} \right)^{-0.43} \left( \frac{M_p }{M_{p,J}} \right)^{1.38} \label{eq:scalingfc} 
%f_{\rm cyc} &\sim &  28~{\rm MHz} \left( \frac{t}{4.5~\rm{Gyr}} \right)^{-0.43} \left( \frac{M_p }{M_{p,J}} \right)^{1.04} \label{eq:scalingfc} 
%\end{eqnarray}
%%%%%%%%%%
%under this model. 
Note that the cyclotron frequency of Jovian planets typically falls between 10~MHz, the plasma cut-off frequency of Earth's ionosphere, and less than 1~GHz. 
In this regime, there are a number of current and near-future radio wave detectors including GMRT, 
LOFAR, Hera, SKA, and upgrades to the VLA.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htbp]
   \plotoneh{fig1_linear.pdf}
   \caption{Planetary magnetic field as a function of age and planetary mass, based on the scaling of magnetic field of \citet{christensen2010} and the evolution of internal heat of \citet{burrows_et_al2001}.} %\memoDS{Can we change density to cgs in this figure?}
  \label{fig:planetaryB_pre}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htbp]
   \plotoneh{fig2_Christensen.pdf}
   \caption{Planetary magnetic field as a function of age for different planetary mass, based on the scaling of magnetic field of \citet{christensen2010} and the evolution of luminosity of \citet{burrows_et_al2001}.} %\memoDS{Can we change density to cgs in this figure?}
  \label{fig:planetaryB}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Estimate of Radio Intensity of RGHJs in comparison with canonical HJs}
\label{ss:brightness}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Based on the assumption that the radio emission is proportional to the input kinetic energy (equation (\ref{eq:Pinp_kin})), the scaling of the radio emission is expanded as follows:
%%%
\begin{eqnarray}
P_{\rm k,\,inp} &=& nv^3 r_s^2 = n v^3 \left( \frac{B^2}{2\pi m_{\rm p}nv^2} \right)^{1/3}  \\
&=& P_{\rm k,\,inp,\,J} \left( \frac{B}{B_J} \right)^{2/3} \left( \frac{a}{5.20~\mbox{AU}} \right)^{-4/3}  \notag \\
&& \times \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \label{eq:Pkinp}
\end{eqnarray}
%%%

We may compare radio emission power of canonical hot Jupiters set at 0.05 AU  and that of RGHJs at 5 AU.
%Canonical values we adopt for planetary mass, orbital distance, and orbital period of each type as shown in Table \ref{tab:comp_HJ}. 
Then, equation (\ref{eq:Pkinp}) can be re-normalized as follows.
%%%
\begin{eqnarray}
P_{\rm k,\,inp} 
&\approx & 227 ~P_{\rm k,\,inp,\,J} \left( \frac{B}{B_J} \right)^{2/3} \left( \frac{a}{5~\mbox{AU}} \right)^{-4/3} \notag \\
&& \times \left( \frac{\dot M_\star}{10^6 \dot M_{\odot}} \right)^{2/3} \left( \frac{v}{10^{-1} v_{\odot}} \right)^{5/3} \\
&& \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \mbox{(for RGB stars' companions)} \notag \\
&\approx & 2.27 \times 10^4 ~P_{\rm k,\,inp,\,J} \left( \frac{B}{B_J} \right)^{2/3} \left( \frac{a}{5~\mbox{AU}} \right)^{-4/3} \notag \\
&& \times \left( \frac{\dot M_\star}{10^9 \dot M_{\odot}} \right)^{2/3} \left( \frac{v}{10^{-1} v_{\odot}} \right)^{5/3}  \\
&& \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \mbox{(for AGB stars' companions)} \notag \\
&\approx & 489 ~P_{\rm k,\,inp,\,J} \left( \frac{B}{B_J} \right)^{2/3} \left( \frac{a}{0.05~\mbox{AU}} \right)^{-4/3} \notag \\
&& \times \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \\
&& \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \mbox{(for canonical hot jupiters)} \notag 
\end{eqnarray}
%%%
Here we normalized the strength of magnetic field of canonical hot jupiters with $B_J$, due to the uncertainty of the magnetic field of tidally-locked planets. 
Depending on the modeling of magnetic field strength, there is also a speculation that the tidally-locked planets may have weaker magnetic field due to slow rotation \citep[e.g.][]{griesmeier2004}; in that case the radio emission would be even weaker. 
%\memoYF{This is based on theoretical prediction by \citet{griesmeier2004} that the tidally-locked planets have weaker magnetic field due to slow spin rotation. }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{table}[htp]
%\caption{Canonical Parameters of 2 types of hot jupiters.}
%\begin{center}
%\begin{tabular}{c|cccc} \hline \hline
%& mass & magnetic field & orbit &  spin period \\ 
%& $M_p$ & B & $a$ &  \\ \hline
%canonical HJ & $M_J$ & $B_J$ & 0.05 AU & 4 days \\
%RGHJ & $M_J$ & $B_J$ & 5 AU &  0.4 days \\ \hline
%\end{tabular}
%\end{center}
%\label{tab:comp_HJ}
%\end{table}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[bp]
   \plotonesc{radio_emission_Mdot1e-8_constRp_christensen.pdf}
   \plotonesc{radio_emission_Mdot1e-5_constRp_christensen.pdf}
   \caption{[{\bf New version: based on scaling relationship of Christensen 2010}] Top panel: Intensity of radio emission (in the unit of Jy) from a companion to a RG star (left) and an AGB star at 10 pc.
The hashed region with vertical lines are not observable because of the plasma frequency cut-off of Earth's ionosphere. 
he hashed region with horizontal lines are not observable because of the plasma frequency cut-off of the stellar wind plasma in the vicinity of the companion.
Bottom panel: cyclotron frequency, i.e., the frequency of radio emission. }
  \label{fig:radio}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Combining the expressions above, we find that the radio spectral flux density observed at the Earth is:
%%%
\begin{eqnarray}
F_{\nu} &=& \frac{P_{\rm radio}}{\Omega d^2 f_{\rm cyc}} \\
&\approx & 5.22\times 10^{-6} \mbox{Jy} \left( \frac{d}{10\mbox{~pc}} \right)^{-2}  \notag \\
&&\times \left( \frac{B}{B_J} \right)^{-1/3}  \left( \frac{a}{5~\mbox{AU}} \right)^{-4/3} \notag \\
&& \times \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \label{eq:F_nu} \\
&& \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \mbox{(for Jupier-twin)} \notag \\
%%%
&\approx & 1.12 \times 10^{-3} \mbox{Jy} \left( \frac{d}{10\mbox{~pc}} \right)^{-2}  \notag \\
&&\times \left( \frac{B}{B_J} \right)^{-1/3} \left( \frac{a}{5~\mbox{AU}} \right)^{-4/3} \notag \\ 
&& \times \left( \frac{\dot M_\star}{10^6 \dot M_{\odot}} \right)^{2/3} \left( \frac{v}{10^{-1} v_{\odot}} \right)^{5/3} \label{eq:F_nu_RGHJ} \\
&& \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \mbox{(for RGB stars' companions)} \notag \\
%%%
&\approx & 1.12 \times 10^{-1} \mbox{Jy} \left( \frac{d}{10\mbox{~pc}} \right)^{-2}  \notag \\
&&\times \left( \frac{B}{B_J} \right)^{-1/3} \left( \frac{a}{5~\mbox{AU}} \right)^{-4/3} \notag \\ 
&& \times \left( \frac{\dot M_\star}{10^9 \dot M_{\odot}} \right)^{2/3} \left( \frac{v}{10^{-1} v_{\odot}} \right)^{5/3} \label{eq:F_nu_AGB} \\
&& \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \mbox{(for AGB stars' companions)} \notag \\
%%%
&\approx & 2.42 \times 10^{-3} \mbox{Jy} \left( \frac{d}{10\mbox{~pc}} \right)^{-2}  \notag \\
&&\times \left( \frac{B}{B_J} \right)^{-1/3} \left( \frac{a}{0.05~\mbox{AU}} \right)^{-4/3} \notag \\ 
&& \times \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \label{eq:F_nu_RGHJs} \\
&& \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \mbox{(for canonical hot jupiters)} \notag 
\end{eqnarray}
%%% 
%\memoYF{cite original paper!}. 
Thus, RGHJs are expected to be intrinsically as bright as the closest hot Jupiters, and Jupiter-like planets around AGB stars are potentially orders-of-magnitude brighter than them. 
Note that, within this framework, the {\it spectral} radio flux has  necessarily negative correlation with the planetary magnetic field, as pointed out by \citet{griesmeier2005}.

Using the model for planetary magnetic field strength, we may consider the  radio spectral intensity and maximum frequency of given planetary mass and age. 
Figure \ref{fig:radio} shows the contours of spectral radio flux of planetary companions at 4.5~Gyr with varying masses and orbital distances. 
The target system is located at a distance of 10~pc from the Earth; the flux can be scaled by the distance as a quadratic function. 
The spectral radio emission from RGHJs at around 5-10 AU is found to be $\sim 10^{-6}-10^{-3}$ Jy when observed from 10-100 pc away.

Observable energy flux is, however, limited by the plasma cut-off frequency in the vicinity of the planets, given by Eq.~(\ref{eq:fplasmalos_scaled}).
We can see emission only from where the cyclotron frequency (equation (\ref{eq:fcyc})) exceeds the plasma frequency:
%%%
\begin{equation}
\label{eq:fcyc>fplasma} f_{\rm cyc} > f_{\rm plasma}^{\rm los}
\end{equation} 
%%%
The hashed regions in Figure \ref{fig:radio} indicate the parameter spaces where this criteria is not satisfied and thus the radio emission from the planet cannot be observed from the Earth.

Equation (\ref{eq:fcyc>fplasma}) can also be rewritten as: 
%%%
\begin{eqnarray}
 \left( \frac{B}{10~\mbox{G}} \right) &>& 0.526 \left( \frac{a}{5~\mbox{AU}} \right)^{-1} \\
 && \times \left( \frac{\dot M_\star}{10^6 \dot M_{\odot}} \right)^{1/2}  \left( \frac{v}{10^{-1}v_{\odot}} \right)^{-1/2} \label{eq:Ba_limit}
\end{eqnarray}
%%%
or, 
%%%
\begin{eqnarray}
 \left( \frac{M_p}{M_{p,J}} \right) &> & 0.627 \left( \frac{a}{5~\mbox{AU}} \right)^{-0.72} \left( \frac{t}{4.5~\mbox{Gyr}} \right)^{0.31} \\
 && \times \left( \frac{\dot M_\star}{10^6 \dot M_{\odot}} \right)^{0.36}  \left( \frac{v}{10^{-1}v_{\odot}} \right)^{-0.36} \label{eq:Ma_limit}
\end{eqnarray}
%%%
In other words, given any orbital distance, there is a lower mass limit for a planet to be able to release the  radio wave out of the system. 
Substituting equation (\ref{eq:Ba_limit}) into equation (\ref{eq:F_nu_RGHJ}) leads to the maximum spectral radio flux at given orbital period:
%%%
\begin{eqnarray}
F_{\nu} &<& 1.30 \times 10^{-3}~\mbox{Jy} \left( \frac{d}{10~\mbox{pc}} \right)^{-2} \\
&& \times \left( \frac{a}{5~\mbox{AU}} \right)^{1/3}  \left( \frac{\dot M_\star}{10^6 \dot M_{\odot}} \right)^{1/2} \left( \frac{v}{10^{-1} v_{\odot}} \right)^{13/6}
\end{eqnarray}
%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure*}[bp]
 %\begin{minipage}{0.5\hsize}
  %\begin{center}
   %\includegraphics[width=\hsize]{radio_emission_Mdot1e-8_constRp.pdf}
%   \plotonesc{radio_emission_Mdot1e-8_constRp.eps}
  %\end{center}
 %\end{minipage}
 %\begin{minipage}{0.5\hsize}
   %\begin{center}
   %\includegraphics[width=\hsize]{radio_emission_Mdot1e-5_constRp.pdf}
%   \plotonesc{radio_emission_Mdot1e-5_constRp.eps}
  %\end{center} 
 %\end{minipage}
%   \caption{[{\bf Old version: based on scaling relationship of Sano 1993}] Top panel: Intensity of radio emission (in the unit of Jy) from a companion to a RG star (left) and an AGB star at 1 pc. The hashed region with vertical lines are not observable because of the plasma frequency cut-off of Earth's ionosphere. The hashed region with horizontal lines are not observable because of the plasma frequency cut-off of the stellar wind plasma in the vicinity of the companion. Bottom panel: cyclotron frequency, i.e., the frequency of radio emission. }
%  \label{fig:radio}
%\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Observability}
\label{s:observability}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this section, we discuss the observability of the estimated radio emission. 
An obvious obstacle is the intrinsic radio emission from host red giant stars themselves; if they are larger than radio flux from the planets, there is no way to identify the latter.
We will see in \S\ref{ss:RGradio} that radio flux from the planets will probably be larger in a certain parameter range. 
Then, we estimate the radio emission of Jupiter-twin around known M giants in \S\ref{ss:actualMgiants}. 
In \S\ref{ss:detectability}, we estimate the sensitivities and limitations of several current and future
radio instruments at lower frequencies.
Finally, \S\ref{ss:timevariability} discusses the expected time variability of the radio emission as a mean to discern the signals from the background noise.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Intrinsic radio emission of red giants stars}
\label{ss:RGradio}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The radio intensity of evolved stars in the range of 10~MHz and 1~GHz is not constrained through observation. 
Above 1~GHz, there are several observations.
For example, \citet{gorman2013} reported $\alpha $ Boo ($R_\star = 25.4 R_{\odot }$, $T=4286$ K, $d=11.26$ pc) at 1~GHz is about 70 $\mu$Jy, and $\alpha $ Tau ($R_\star = 44.2  R_{\odot }$, $T=3,910$ K, $d=20.0$ pc) at 3~GHz is about 40 $\mu$Jy.

We estimate the thermal radio flux from the host star to be consistent with these observed data and by extrapolating to the lower frequency range with a simple power law:
%%%
\begin{eqnarray}
\nonumber f_{\star }(\nu ) &\sim & (0.4-1.375) \times 10^{-3} ~\mbox{Jy} \\
&& \times \left( \frac{d}{10 ~\mbox{pc}} \right)^{-2}  \left( \frac{R_{\star }}{100 R_{\odot }} \right)^2 \left( \frac{\nu}{1~\mbox{GHz}} \right)^{\alpha } 
\end{eqnarray}
%%%
Constant property wind model \citep{wright1975,panagia1975,olnon1975} predicts $\alpha = 0.6$, while observations suggested that the power index $\alpha $ is steeper than 0.6 \citep{gorman2013}.

Figure \ref{fig:cartoon} is a cartoon for a radio spectrum of a planet with $10~M_J$ and that of the host red giant stars with 100~$R_{\odot }$, both of which is placed at 10 pc away. 
As this figure indicates, even if the radio emission from the red giant itself has a relatively shallow spectral index, the radio emission from the RGHJ still dominates in the 10-200~MHz range. 
Therefore, in the following, we ignore the radio emission from the red giant as a noise source.

% Neil's addition

Over the last four decades there have been numerous radio continuum observations of nearby red giants, many with the aim of understanding the extended atmospheres and mass-loss mechanisms of K- and M-type giants~\citep[e.g.,][]{Newell1982, Knapp1995, Skinner1997, Lim1998, OGorman2013}.
Almost all observations have been carried out at wavelengths of 6 cm and below, probing thermal Bremsstrahlung emission from the large, partially ionized envelopes surrounding the giant stars~\citep{drake1986}.
The spectral index of this emission is of order unity, with a range of reported values for various sources between 0.8 and 1.6~\citep{OGorman2013}.

Only two published studies describe attempts to detect continuum emission from single (non-binary) red giant stars at wavelengths above 20 cm ($L$ band); both yielded null results.
First, a program that observed a sample of nine M-type giants at 430 MHz with the Arecibo Telescope failed to detect any of the sources down to flux density 10 mJy~\citep{Fix1976}.
Later, with the Molonglo Observatory Synthesis Telescope (MOST), a sample of eight K- and M-type giants were observed at 843 MHz; the upper limits of their flux densities were placed at approximately 1 mJy~\citep{Beasley1992}.
Current facilities could achieve far deeper sensitivity on similar targets.
For example, as listed in Table~\ref{tab:sens}, the Giant Metrewave Radio Telescope (GMRT) at 150 MHz reaches a 
sensitivity better than 0.5~mJy in under an hour.
However, any re-attempts to detect single red giants at wavelengths near 1 meter remain to be presented in the literature.

The only red giants that have been detected at meter wavelengths are those in interacting binary systems like those of the RS CVn type.
For example, van den Oord and de Bruyn detected plasma maser emission from II Pegasi at 360 MHz and 609 MHz with the WSRT~\citep{vandenOord1994}.
Although single dwarf-type main sequence stars are also known to exhibit bright, coherent radio flares~\citep{Bastian1990}, no analog to this non-thermal emission process has been proposed to exist for evolved stars.

Altogether, we lack firm observational constraints on the brightnesses of M giants near 1-m wavelengths.
However, assuming thermal Bremsstrahlung emission continues to dominate in this wavelength regime, we can extrapolate down from the reported centimeter flux densities, using the specific spectral index when available.
In this case, for example, extrapolating from the 0.4 mJy flux density of $\alpha$ Ori at 1.46 GHz, and using its spectral index 1.32~\citep{Newell1982}, we arrive at an estimated flux density of 0.04 mJy at 300 MHz.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tbp]
   \plotonesc{cartoon_10Mj_2.pdf}
   \caption{A cartoon of radio emission spectra of a RGHJ with 10$M_J$ and the host red giants with 100 $R_{\odot }$.
The spectrum of RGHJ is modeled after Jovian radio spectra, e.g. figure 8 of \citet{zarka_et_al2004} and figure 2 of \citet{griesmeier2007a}; contribution from Io is not shown here.
The spectra of the host red giant are modeled simply by extrapolating observed radio spectra above 1~GHz with power law. }
  \label{fig:cartoon}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Radio Flux from Nearby M giants}
\label{ss:actualMgiants}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[tbhp]
   \plotoneh{radio_1Mp.pdf}
   \plotoneh{radio_10Mp.pdf}
%   \plotoneh{radio_1Mp_5AU.pdf}
%   \plotoneh{radio_5Mp_5AU.pdf}
%   \plotoneh{radio_2Mp_4-5Gyr.pdf}
%   \plotoneh{radio_4Mp_4-5Gyr.pdf}   
   \caption{Radio emission from existing M-type red giants with a hypothetical planetary companion of 1$M_J$ and 10$M_J$ at 1 AU and 5AU.
Cross symbols indicate that the radio emission is not observable because cyclotron frequency is less than plasma density around the planet. }
  \label{fig:observability}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In order to examine the potential to detect radio emission of RGHJs with the current and near-future instruments, we estimate the possible radio emission from known red giants assuming that they have planetary companions. 
For this purpose, we obtained the list of M-type red giants from Table 4 of \citet{dumm1998} with the data of mass, radius, and effective temperature.
For each red giant, the mass-loss rate is estimated by Reimers' equation \citep{reimers1975}:
%%%
\begin{equation}
\dot M_\star [M_\odot{\rm /yr}] \sim 4 \times 10^{-13} \eta \left( \frac{R_{\star }}{R_{\odot }} \right)\left( \frac{L_{\star }}{L_{\odot }} \right) \left( \frac{M_\star}{M_\odot} \right)^{-1}
\end{equation}
%%%
where $R_{\star }$, $L_{\star } (=4\pi R_{\star }^2 T_{\star }^4)$, and $M_{\star}$ are radius, luminosity, and mass of the star. 
The factor $\eta $ is $\sim 0.3-3$, and we assume $\eta =1$ here. 
The velocity of stellar wind is assumed to scale with the escape velocity, hence
%%%
\begin{equation}
v \propto \sqrt{\frac{2GM_\star}{R_{\star }}}
\end{equation}
%%%
The age of the system is assumed to roughly same as the lifetime of their main-sequence stage, which depends on the stellar mass.
We assume \memoYF{validation}:
%%%
\begin{equation}
t \sim 10\,{\rm Gyr}\, \left( \frac{M_{\star}}{M_{\odot }} \right)^{-2.5}
\end{equation}
%%%
In addition, the distance to the system is obtained based on the parallax data from Hipparcos datasets\footnote{http://www.rssd.esa.int/index.php?project=HIPPARCOS}.

Using these data, we can now calculate the maximum frequency and intensity of radio emission by specifying planetary mass and orbital distance of a hypothetical planetary companion. 
We consider [A] a Jupiter-twin whose cyclotron frequency is $\sim 30$~MHz, and [B] a larger Jovian planet with $M_p\sim 10M_p$ whose cyclotron frequency is $\sim 300$~MHz (equations (\ref{eq:scalingB}) and (\ref{eq:fcyc})).

Figure \ref{fig:observability} displays the estimated radio intensity (i.e., cyclotron frequency) of Jupiters [A] and [B] around M-type red giants within 100~pc. 
Jupiters are placed at 1 AU and at 5 AU for reference. 
The difference of the symbols indicates whether the radio emission can escape from the system (circle) or not (cross).

%\memoYF{Up to where would the planet be engulfed by evolved stars? References?}
%I wondered this because if planets at 1 AU would be engulfed anyway it would not be worth discussing here. According to e.g. Nordhaus and Spiegel (2013) it is the case--however, some planets may migrate inward within 1 AU after evolution. So it should be ok. In any case this is not the main point of this paper and I would leave it out.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Detectability with Current and Near Future Low Frequency Radio Instrumentation}
\label{ss:detectability}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{deluxetable*}{cccccc}
\tabletypesize{\footnotesize}
\tablecolumns{6}
\tablecaption{Detectability of Hot Jupiters with Current 
and Future Radio Telescopes.\label{tab:sens}}
\tablehead{
Instrument & Band  & \sigRMS\tablenotemark{a} & $\tau$\tablenotemark{b}  & Resolution & $\sigma_c$\tablenotemark{c}\\
           & (MHz) & ($\mu$Jy~hr$^{1/2}$)         &  (hr)                    & (\arcsec)  & ($\mu$Jy bm$^{-1}$)} 
\startdata
LOFAR-LBA  &  30-80  & 700   & 120000             & 8.3  & 42 \\ 
LOFAR-HBA  & 110-200 &  35   &    310             & 3.3  & 1.1 \\  
HERA       &  30-80  & 870   & 160000             & 1030 & 4.2$\times10^5$\\ 
GMRT       & 130-190 & 430   &  46000             & 16.5 & 56.9 \\ 
%LWA       &  30-80  & 58000 &  702127141         & 10.3 & 42.2 \\
%MWA       &  30-80  & 29000 &  176223211         & 1030 & 4.2$\times10^5$ \\
SKA-Low (part) & 50-200 & 7.8 &  12.6             & 7.6  & 9.7 \\
SKA-Low (full) & 50-350 & 2.2 &  1.0              & 4.8  & 1.5 \\
%VLA 4-Band& 30-80   & 3300  & 2312058            & 28.6 & 330 \\
VLA-LOBO   &  50-350 &  92   &  1800              & 8.6  & 10.4
\enddata
%\tablecomments{Time estimates to reach 2~$\mu$Jy RMS.}
\tablenotetext{a}{RMS noise in a 1-hour observation, computed for imaging of the full band reported 
in this table (Equation~\ref{eq:rms_sens}).}
\tablenotetext{b}{The integration time in hours required to reach an RMS of 2~$\mu$Jy (i.e.\ 5-$\sigma$ detection
of a typical source at 100~pc with 10~$\mu$Jy bm$^{-1}$ average flux).}
\tablenotetext{c}{Source confusion RMS contribution.}
\end{deluxetable*}
%
% HERA
% http://reionization.org/wp-content/uploads/2015/05/HERA0_imaging.pdf
%
% LOFAR
% https://www.astron.nl/~heald/test/sens.php; 
% 24 core stations, 16 remote, 8 international, 488 subbands
%
% GMRT
% http://gmrt.ncra.tifr.res.in/~astrosupp/obs_setup/sensitivity.html
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We consider in this section several current and near future radio observatories that operate at frequencies $\nu \lesssim 350~\rm MHz$ and can reach continuum sensitivities less than 1~mJy in a few hours.
The results are summarized in Table~\ref{tab:sens}.

The primary limitation to detectability is a telescope's sensitivity, characterized by the root mean square (RMS) noise fluctuations in the sky and receiver.
The RMS noise \sigRMS\ in an interferometric observation with integration time $\tau$, bandwidth $\Delta \nu$, effective area $A_{\rm eff}$ ($\approx0.7 \pi (D/2)^2$ for interferometers comprising antennas with dish size $D$), and $N$ antennas is \citep[see e.g.][]{condon+ransom2016}
%%%
\begin{equation}
\sigRMS = \frac{2 k T_{\rm sys}}{A_{\rm eff} \, \sqrt{\Delta \nu \, \tau N (N-1)}} 
\left(\frac{\rm 10^{26} Jy}{\rm W m^{-2} Hz^{-1}}\right).
\label{eq:rms_sens}
\end{equation}
%%%
Here, \Tsys\ is the blackbody temperature equivalent of the total system noise, which is the sum of the instrument or receiver noise temperature \Trx\ and the noise contribution from the sky \Tsky.
We estimate \Tsky\ using the low frequency sky noise temperature fit from \cite{Rogers+Bowman2008}:
%%%
\begin{equation}
\Tsky = T_{150} \left( \frac{\nu}{150~\rm MHz}\right)^{-\beta} + \Tcmb 
\end{equation}
%%%
where $T_{150} = 283.2~\rm K$ is the \Tsky\ at 150~MHz, $\beta=2.47$, and $\Tcmb = 2.73~\rm K$ is the contribution from the cosmic microwave background (CMB).
While this relation was fit to data taken in the 100-200~MHz range, the authors found it to be consistent with published measurements from 10-408~MHz. 
%The \Trx\ is the receiver noise temperature, which depends on the instruments and is described below.

An additional contribution to the noise is due to source confusion, the imaging noise associated with unresolved interloping radio sources in an observation.
Source confusion is characterized by the standard deviation $\sigma_c$ in the surface brightness of an image due to one or more unresolved sources in the beam solid angle (for a review, see \cite{Condon1974,Condon2012}).
To estimate $\sigma_c$, reported in Table~\ref{tab:sens}, we use the relations provided by \cite{condon+ransom2016}, reproduced below:
%%%
\begin{eqnarray}
\sigma_c & \approx & 0.2 \left( \frac{\nu}{1 \rm~GHz} \right)^{-0.7} \left( \frac{\theta}{1\arcmin} \right)^2\!\!\!\!,~(\theta > 10\arcsec) \\
\sigma_c & \approx & 2.2 \left( \frac{\nu}{1 \rm~GHz} \right)^{-0.7} \left( \frac{\theta}{1\arcmin} \right)^{10/3}\!\!\!\!\!\!\!\!\!\!,~(\theta < 10\arcsec).
\end{eqnarray}
%%%
The source confusion limit is dominated by active galactic nuclei, which have low polarization fractions $\lesssim2.5\%$ \citep{Stil2014} and slowly vary with time.
Therefore both the circularly polarized nature of the emission from hot Jupiters and the time variability (discussed in \S\ref{ss:timevariability}) partially mitigate the source confusion limit, allowing hot jupiters, like other time-variable sources, to be detected below the imaging confusion limit.
We therefore expect confusion not to be a strong limitation to detectability of hot Jupiters.

% LOFAR
The LOw-Frequency ARray (LOFAR) operates as two separate arrays defined by their high and low bands, differing configurations, and antenna designs, which are called respectively the Low Band Antennas (LBA) and High Band Antennas (HBA).
LOFAR-LBA operates in the 10-90~MHz range, and LOFAR-HBA operates in the 100-240~MHz range \citep{vanHaarlem2013}.
We consider it as two separate instruments here.
LOFAR-LBA's most sensitive band is centered at 60~MHz, and LOFAR-HBA's is at 150~MHz.
Given the strong frequency dependence of the effective aperture of a dipole antenna, and that the reciever noise temperature for LOFAR $\Trx \sim \Tsky$, we use the LOFAR Image Noise calculator\footnote{Heald; \url{http://www.astron.nl/~heald/test/sens.php}}, which is based on {\it SKA Memo 113}\footnote{Nijboer, Pandey-Pommier, \& de Bruyn; \url{http://www.skatelescope.org/uploaded/59513\_113\_Memo\_Nijboer.pdf}}, for our fiducial sensitivity and imaging parameters in Table~\ref{tab:sens}.

% HERA
The Hydrogen Epoch of Reionization Array (HERA\footnote{\url{http://reionization.org/}}) \ldots

% GMRT
The Giant Metre-wave Radio Telescope (GMRT) is an array comprising thirty 25-meter antennas, operating at frequencies 130-2000~MHz.
The lowest band (130-190~MHz) exhibits receiver noise $T_{\rm rx}$ comparable to the sky noise.
To be conservative, we use the GMRT online calculator.\footnote{\url{http://gmrt.ncra.tifr.res.in/~astrosupp/obs\_setup/sensitivity.html}}

% SKA
The low frequency component of the Square Kilometer Array (SKA), recently rebaselined to have 0.4~km$^2$ collecting area in 130,000 elements with baselines extending to 65~km, will operate from 50-350~MHz.\footnote{\url{https://www.skatelescope.org/news/worlds-largest-radio-telescope-near-construction/}}
We estimate the noise in the lower half of the band ($\nu<200~\rm MHz$), appropriate for lower mass ($\lesssim 10 M_J$) separately from the full band, which could be used to detect $\sim 30 M_J$ hot Jupiters.
We adopt the \cite{Rogers+Bowman2008} relation for \Tsys\ as above, and assume $\Trx \approx 60~\rm K$.
We caution that these assumptions are optimistic, and the time estimates could in reality be a factor of a few worse.

% VLA
For the low frequency receivers (28-80~MHz; 4-Band) in the \memoDS{Long Wavelength Array (LWA) and the?}\memoTM{No need to mention LWA here.  The receivers are the same in the VLA 4-band and the LWA, but the LWA is far too insensitive to consider here, as is VLA 4-band alone, so the reference just serves to place an upper limit on what \Trx\ will be for LOBO.} Very Large Array (VLA), the receiver noise temperature $\Trx \approx 260~\rm K$ \citep{Hicks2012} and is subdominant to \Tsky\ by at least -6~dB.
Future upgrades to the VLA such as the LOw Band Observatory \citep[LOBO][]{Kassim2015IAU} are being considered, and would cover the full 30-470~MHz band.
We estimate the sensitivity of LOBO assuming the receiver noise extrapolates linearly from $\Trx \approx 60~\rm K$ (internal VLA memo) in P-Band (230-470~MHz) to $\Trx \approx 260~\rm K$ in 4-band, and consider just the 50-350~MHz portion of the full 30-470~MHz band.

The estimates of the noises shown in Table \ref{tab:sens}, together with Figure \ref{fig:observability}, indicate that, at the frequency relevant to the radio emission from Jupiter-size planets, current instrument sensitivities are too low for detection.
On the other hand, for more massive exoplanets ($M_p\sim 10 M_J$) where the bulk of the radio emission would exist above 50~MHz, there is hope to detect radio emission from nearby systems with LOFAR-HBA or LOBO.
LOFAR-HBA has reasonable sensitivity to $M_p =10~M_J$ planets at 5 AU around a few closest systems as well as those around further stars but orbiting closer to the stars. 
More promising is the low frequency component of SKA, which will be able to detect the signal within one day from planets with $M_p \sim 10~M_J$ and orbital distances $\sim$5~AU around M stars $<$100~pc from Earth.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Discriminating from Background:\\Time Variability}
\label{ss:timevariability}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%In order to identify radio emission from exoplanets in the galactic background noise, it is

In order to identify radio emission from exoplanets, it is critical to discern it from the galactic background. 
An important feature useful for this purpose is the time variability of radio emission. 
Indeed, several processes cause the time variability of radio emission from exoplanets at different timescales; they are listed below from the ones typically with shorter timescales to those with longer timescales.

\paragraph{Modulation due to planetary spin rotation (a few hours $\sim $)} 
If the planetary spin axis is misaligned to the planetary magnetic axis, like Jupiter, the radio intensity will display sinusoidal modulation synchronously to the spin period.
%

\paragraph{Modulation due to the presence of satellites (a few hours $\sim $)}

\paragraph{Variability of stellar wind due to stellar oscillation (day $\sim $)?}
Based on the correlation between the planetary radio emission and the solar wind plasma density also suggest that the radio emission from the planet around the evolved stars, which typically oscillate, are time variable.
\memoYF{To Jason: time variability?}
%
\paragraph{Modulation due to orbital motion of the planet (a few years $\sim $)}
If the planetary orbit is close to the edge-on configuration, the radio emission disappears synchronously to the orbital period, due to the increased plasma cut-off frequency along the path and/or secondary eclipse.
The former is because when the planet is behind the star the path of the radio emission toward the Earth comes through the vicinity of the star where the plasma density is high.
The probability of the latter one, secondary eclipse, is $\sim a/R_\star \sim 9$\% (in the case that $a=5~\mathrm{AU}$ and $R_{\star}=100R_{\odot }$). 
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
\label{s:discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Back-reaction of a plasma flowing into a magnetic field}
\label{ss:offset}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

One might expect that plasma flowing into a region permeated by a magnetic field would, by spiraling around the field lines, generate an opposing magnetic field that partially cancels the intrinsic planetary magnetic field. 
To estimate the strength of this effect, consider a uniform flow of particles of charge $e$, mass $m$, and velocity $v$ into a region of magnetic field $B$, spiraling along the initial magnetic field. 
A single charged particle moving perpendicular to the magnetic field moves in a circular orbit of radius 
%%%
\begin{equation}
r=\frac{mv_\bot }{eB}.
\end{equation}
%%%
which creates a magnetic dipole
%%%
\begin{equation}
\mu = \frac{e}{2\pi r/v_\bot} \pi r^2 = \frac{1}{2} e r v_\bot = \frac{mv_\bot^2}{2B}.
\end{equation}
%%%
The volume integral of the canceling magnetic field $B_c$ generated by a single dipole $\mu$ is
%%%
\begin{equation}
\int d^3{\boldsymbol r} B_c = \frac{8\pi}{3} \mu.
\end{equation}
%%%

Hence, for a number density $n$ of these charged particles, the fraction of canceling field to the background magnetic field is
%%%
\begin{equation}
\frac{B_c}{B} = \frac{n(mv_\bot^2/2)}{3 B^2/8\pi} < \frac{\rho_K}{3\rho_B},
\end{equation}
%%%
where $\rho_K$ and $\rho_B$ are the kinetic energy density and the magnetic energy density, respectively. 
%If the initial velocity is not perpendicular to the magnetic field, $v$ must everywhere be replaced by its perpendicular component.

Just inside the stand-off point, 
%%%
\begin{equation}
m_p n[r_s] v^2 < \frac{B[r_s]^2}{2\pi }
\end{equation}
%%%
and thus $\rho_K/\rho _B < 1$. 
When the particle spirals into the planet, 
the magnetic pressures increases (by beaming?) by 
%%%
\begin{equation}
\frac{B^2}{8\pi } \propto \frac{1}{r^6}
\end{equation}
%%%
On the other hand, the kinetic energy of the moment increases
%%%
\begin{equation}
\frac{mv^2}{2} \propto \frac{1}{r}
%\rho _K \propto \frac{1}{r}
\end{equation}
%%%
Therefore, unless the density increases more drastically than $1/r^5$, the ratio $B_c/B$ is significantly less than unity.

%It is interesting to note that, for a plasma of protons and electrons in a stellar wind, the canceling field is clearly dominated by the contribution of protons, though the cyclotron emission is dominated by electrons.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
\label{sec:conc}

In this paper, we propose that ``hot Jupiters'' around evolved stars (RGHJ) may be bright radio wave emitters, on the assumption of the simple empirical correlation between the radio emission and the kinetic input energy. 
Massive stellar wind of red giants and AGB stars would put the kinetic energy into magnetospheres of distant Jovian planets, which could lead to the intrinsic bright radio emission, comparable to or greater than canonical hot Jupiters in close-in orbits.

The maximum radio emission we can expect to observe is limited by the plasma cut-off frequency of the stellar wind in the vicinity of the planets. 
For a planet with the same magnetic moment as Jupiter, the maximum radio emission from the planets which can penetrate through the surrounding plasma is $\sim 10^{-3}$ Jy or $\sim 10^{-5}$ Jy at 10~pc and 100~pc, respectively.
As we showed in \S\ref{ss:detectability}, the upper end of this range is detectable with the current generation of radio facilities, while farther hot Jupiters should soon be detectable using the next generation of instruments.

Admittedly, uncertainties exist in the relation between radio emission of Jovian planets and the input energy from stellar wind, as well as in the scaling of planetary magnetic field. 
Alternatively, once we detect both the gravitational perturbation and radio emission of Jovian planets, these models will be contested.
The viability of these scaling relations will be examined in the future.

%The planetary magnetic moment is computed based on the simple scaling relationship given by \citep{christensen2010}. 
%-about detectability
%-number of evolved stars within 100pc $\sim 1000$.

\vspace{0.5in}

\acknowledgements

{\sc Acknowledgments}

YF is supported from the Grant-in-Aid No. 25887024 by the Japan Society for the Promotion of Science.
DSS gratefully acknowledges support from a fellowship from the AMIAS group.
We thank David Hogg for encouraging us to pursue calculations of explanetary radio emission.
NM acknowledges support from [???].
The portion of this research for which TM is responsible was performed while he held a National Research Council Research Associateship Award at the Naval Research Laboratory (NRL).
Basic research in radio astronomy at NRL is supported by 6.1 Base funding.

\bibliography{biblio.bib}

\newpage

\appendix

\section{Comparison with previous work}

There are rather wide variety of estimation for radio emission of Jovian exoplanets and it is worth noting how the different models return different answers.

A common part is the scaling relationship between the planetary magnetic field strength and the input kinetic energy.
Combining equation (\ref{eq:Pinp_kin}), i.e.,
%%%
\begin{equation}
P_{\rm inp,k}=nv^3 r_s^2 \propto \dot M_{\star} v^2 a^{-2} r_s^2 s
\end{equation}
%%%
and the scaling between the radius of the magnetic stand-off point, $r_s$, and the planetary magnetic field $B$ (see equation (\ref{eq:stand-off-radius})):
%%%
\begin{equation}
r_s \propto B^{1/3} a^{1/3} \dot M_{\star }^{-1/6} \dot v^{-1/6}
\end{equation}
%%%
under the constant stellar wind, we obtain:
%%%
\begin{equation}
P_{\rm inp, k} \propto B^{2/3} a^{-4/3} \dot M_{\star }^{2/3} \dot v^{5/3}
\end{equation}
%%%

%equation (\ref{eq:stand-off-radius})
%%%
%\begin{equation}
%r_s = 11.3 \, R_J \left( \frac{B}{10~\mbox{G}} \right)^{1/3}  \left( \frac{a}{5\mbox{AU}} \right)^{1/3} \left( \frac{\dot M_\star}{10^6 \dot M_\odot} \right)^{-1/6}  \left( \frac{\dot v}{10^{-1}v_{\odot }} \right)^{-1/6}
%\end{equation}
%%%
Aside from this common part, the differences originate mainly from:
\renewcommand{\theenumi}{\roman{enumi}}
%%%
\begin{enumerate}
\item Assumptions for the scaling law of radio power in terms of input energy. 
\item Assumptions for the scaling law of planetary magnetic field strength. 
\item Different normalization (assumptions for Jovian radio emission). 
\item Assumptions for band width (either $f_{\rm cyc}$ or $0.5f_{\rm cyc}$). 
\item Modeling of stellar wind
\end{enumerate}
%%%

As for i., ii. and iii., the scaling laws are summarized in Table \ref{tab:comp}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}[htp]
\caption{Comparison for assumed scaling relationships.}
\begin{center}
\begin{tabular}{c|l|llcc} \hline \hline
%
& Paper & Radio power & Magnetic moment  & norm [W] & band width \\\hline
%
\#1 & \citet{farrell1999} & $P_{\rm radio} \propto P_{\rm sw}^{0.88} $ & $\mathcal{M} \propto \omega M_p^{5/3}$ & $4 \times 10^9$ & $ 0.5 f_{\rm cyc}$ \\ \hline
%
\#2 & \citet{farrell1999} & $P_{\rm radio} \propto P_{\rm sw}^{1.2} $ & $\mathcal{M} \propto \omega M_p^{5/3}$ & $4 \times 10^{11}$ & $ 0.5 f_{\rm cyc}$ \\
& \citet{lazio2004} & & $\mathcal{M} \propto \omega M_p^{2}$?? & & \\
& \citet{ignace2010} & & & & \\ \hline
\#3 & \citet{stevens2005} & $P_{\rm radio} \propto P_{\rm sw} $ & $\mathcal{M} \propto M_p$ & - & - \\ \hline
\#4 & \citet{griesmeier2005,griesmeier2007a,griesmeier2007b} & $P_{\rm radio} \propto P_{\rm sw}$ & $\mathcal{M} \propto \rho _c^{1/2} \omega r_c^4 $ & $2.1 \times 10^{11}$ & $ 0.5 f_{\rm cyc}$ \\
\#5 & & & $\mathcal{M} \propto \rho _c^{1/2} \omega^{1/2} r_c^3 \sigma ^{-1/2} $ & & \\
\#6 & & & $\mathcal{M} \propto \rho _c^{1/2} \omega^{3/4} r_c^{7/2} \sigma ^{-1/4} $ & \\
\#7 & & & $\mathcal{M} \propto \rho _c^{1/2} \omega r_c^{7/2}$ & & \\ \hline
%\#8 & \citet{griesmeier2007a,griesmeier2007b} & $P_{\rm radio} \propto P_{\rm sw}$ & same as above & $2.1 \times 10^{11}$ & $ f_{\rm cyc}$ \\ \hline
\#8 & \citet{reiners2010} & $P_{\rm radio} \propto P_{\rm sw}$ & $\mathcal{M} \propto M_p^{1/6} L^{1/3} R_p^{25/6}$ \\ 
 & this work & & \\ \hline
\end{tabular}
\end{center}
\label{tab:comp}
\tablecomments{}
\end{table*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

If we use the assumptions that $R_p \sim R_J\;\;(const.)$, $\rho _c\sim R_p\sim const.$, and $\sigma = const.$, then the assumptions are reduced to:
%%%%%%%%%
\begin{eqnarray}
\mbox{[\#1]} \; P_{\rm rad} &\sim & 4 \times 10^9 \,{\rm W}\,\left( \frac{\omega}{\omega _J} \right)^{0.58} \left( \frac{M_p}{M_J} \right)^{0.98} \left( \frac{a}{a_J} \right)^{-1.17} \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{0.586} \left( \frac{v}{v_{\odot}} \right)^{1.46} \\
%
\mbox{[\#2]} \; P_{\rm rad} &\sim & 4 \times 10^{11} \,{\rm W}\, \left( \frac{\omega}{\omega _J} \right)^{0.79} \left( \frac{M_p}{M_J} \right)^{1.33} \left( \frac{a}{a_J} \right)^{-1.60} \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{0.8} \left( \frac{v}{v_{\odot}} \right)^{2.0} \\
%
\mbox{[\#3]} \; P_{\rm rad} &\sim & ? \times 10^{11} \,{\rm W}\,\left( \frac{M_p}{M_J} \right)^{2/3} \left( \frac{a}{a_J} \right)^{-4/3} \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \\
%
\mbox{[\#4]} \; P_{\rm rad} &\sim & 2.1 \times 10^{11} \,{\rm W}\,\left( \frac{\omega}{\omega _J} \right)^{2/3} \left( \frac{M_p}{M_J} \right)^{1/3} \left( \frac{a}{a_J} \right)^{-4/3} \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \, \\
%
\mbox{[\#5]} \; P_{\rm rad} &\sim & 2.1 \times 10^{11} \,{\rm W}\,\left( \frac{\omega}{\omega _J} \right)^{1/3} \left( \frac{M_p}{M_J} \right)^{1/3} \left( \frac{a}{a_J} \right)^{-4/3} \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \, \\
%
\mbox{[\#6]} \; P_{\rm rad} &\sim & 2.1 \times 10^{11} \,{\rm W}\,\left( \frac{\omega}{\omega _J} \right)^{1/2} \left( \frac{M_p}{M_J} \right)^{1/3} \left( \frac{a}{a_J} \right)^{-4/3} \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \, \\
%
\mbox{[\#7]} \; P_{\rm rad} &\sim & 2.1 \times 10^{11} \,{\rm W}\,\left( \frac{\omega}{\omega _J} \right)^{2/3} \left( \frac{M_p}{M_J} \right)^{1/3} \left( \frac{a}{a_J} \right)^{-4/3} \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \, \\
%
\mbox{[\#8]} \; P_{\rm rad} &\sim & ? \times 10^{11} \,{\rm W}\, \left( \frac{M_p}{M_J} \right)^{0.69}  \left( \frac{t}{4.5 \rm~Gyr} \right)^{-0.288}  \left( \frac{a}{a_J} \right)^{-4/3} \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \, \\
\end{eqnarray}
%%%%%%%%%
For the last equation, we used the scaling of the luminosity $L$, equation (\ref{eq:burrowsLum}):
%%%
\begin{equation}
L \sim 1.84\times10^{-10} L_\odot \left( \frac{t}{4.5 \rm~Gyr} \right)^{-1.3} \left( \frac{M_p}{M_J} \right)^{2.64} \tag{\ref{eq:burrowsLum}}
\end{equation} 
%%%
\memoYF{Note that in \citet{ignace2010} the dependence on mass-loss rate and stellar wind velocity is not taken account consistently; in that paper, $P_{\rm rad} \propto \dot M_{\star } v^2$.}

In summary, we write down the scaling of the radio emission as below, 
%%%
\begin{equation}
P_{\rm rad} \times 10^{11} \,{\rm W}\, \left( \frac{\omega}{\omega _J} \right)^{\alpha } \left( \frac{M_p}{M_J} \right)^{\beta }  \left( \frac{t}{4.5 \rm~Gyr} \right)^{-\gamma }  \left( \frac{a}{a_J} \right)^{-\delta }  \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{\epsilon } \left( \frac{v}{v_{\odot}} \right)^{\phi } \, \\
\end{equation} 
%%%
with $\alpha = 0.5-0.8$, $\beta = 0.33-1.33$, $\gamma = 0-0.288$, $\gamma = 1.17-1.60$, $\epsilon = 0.58-0.8$, $\phi = 1.46-2.0$.

\end{document}

\clearpage

\section{Memo}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Brighter Implies Harder to See:\\
Greater $B$ Produces Lower $F_\nu$}
\label{ss:counterintuitive}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The cyclotron frequency at radius $r$ is:
%%%
\begin{equation}
f_{\rm cyc} = \frac{e B(r)}{2\pi m_e} \propto \frac{B_p}{r^3}
\end{equation}
%%
Therefore, 
%%%
\begin{equation}
\frac{dP}{df_{\rm cyc}} = \frac{dP}{d\nu } \propto -\frac{r^4}{B_p} \frac{dP}{dr}
\end{equation}
%%%
Employing Larmor formula, power emitted from a thin shell at $r\sim r+dr$, $dP/dr$ is 
%%%
\begin{equation}
\frac{dP}{dr} \propto 4\pi r^2 N_{\rm plasma}(r) v_{\bot }^2(r) B^2(r) \label{eq:larmor}
\end{equation}
%%%
where $N_{\rm plasma}(r) $ represents the number density of plasma in the planetary magnetosphere at radius $r$.

We assume that $N_{\rm plasma}(r_s) $  is proportional to the squared radius of the magnetic stand-off point:
%%%
\begin{equation}
N_{\rm plasma}(r_s) \propto \pi r_s^2 \;\;(\propto B_p^{2/3}) \label{eq:N0_plasma}
\end{equation}
%%%
Due the the mass flux conservation,
%%%
\begin{equation}
4 \pi r^2 N_{\rm plasma}(r) v_{\|} = const.
\end{equation}
%%%
Combining with equation (\ref{eq:N0_plasma}),
%%%
\begin{equation}
N_{\rm plasma}(r) \propto  \frac{r_s^2 }{v_{\|}(r) r^2} \propto  B_p^{2/3} \cdot \frac{v_{\|}(r)}{r^2}
\end{equation}
%%%

Because the magnetic dipole moment
%%%
\begin{equation}
\mu = \frac{mv_{\bot}(r)}{eB(r)}
\end{equation}
%%%
is approximately to be constant while the plasma falls down along the magnetic field, 
%%%
\begin{eqnarray}
v_{\bot }(r) \propto \frac{1}{r^3} \label{eq:v_bot} \\
B(r) \propto \frac{B_p}{r^3} 
\end{eqnarray}
%%%
On the other hand, conservation of energy ensures that
%%%
\begin{equation}
\mathcal{E} = -\frac{GM_p}{r} + \frac{v_{\bot }^2}{2} + \frac{v_{\|}^2}{2}=const.
\end{equation}
%%%
Assuming that the initial energy (at $r=r_s$) is negligible compared to those near the surface, $\mathcal{E}=0$, 
%%%
\begin{equation}
v_{\|}^2 \sim \frac{GM_p}{r} - \frac{v_{\bot ,0}^2}{2} \left( \frac{r}{r_s} \right)^{-6} \label{eq:v_||}
\end{equation}
%%%

Substituting equations (\ref{eq:N0_plasma}),  (\ref{eq:v_bot}) and (\ref{eq:v_||}) into equation (\ref{eq:larmor}) leads to
%%%
\begin{equation}
\frac{dP}{dr} \propto \frac{B_p^{8/3}}{r^{12}} \left( \frac{GM_p}{r} - \frac{v_{\bot ,0}^2}{2} \left( \frac{r}{r_s} \right)^{-6} \right)^{1/2}  
\end{equation}
%%%
The right-hand side has the maximum value $\sqrt{5GM_p/6\tilde r} $ at radius $\tilde r=(6v_{\bot ,0}^2/GM_p)^{1/5}$. 
At $r>>\tilde r$, 
%%%
\begin{equation}
\frac{dP}{dr} \propto \frac{B_p^{8/3}}{r^{25/2}}
\end{equation}
%%%
and
%%%
\begin{equation}
\frac{dP}{d\nu } \propto \frac{B_p^{5/3}}{r^{17/2}} \propto \frac{\nu^{17/6}}{B_p^{7/6}}
\end{equation}
%%%

At $r<<\tilde r$, 
%%%
\begin{equation}
\frac{dP}{dr} \propto \frac{B_p^{8/3}}{r^{15}}
\end{equation}
%%%
and
%%%
\begin{equation}
\frac{dP}{d\nu } \propto \frac{B_p^{5/3}}{r^{11}} \propto \frac{\nu^{11}}{B_p^{28/3}}
\end{equation}
%%%

\newpage

\memoDS{Is this section needed anymore, since this is already in the Griesmeier paper?
Does the Griesmeier paper make the point explicit enough, or is this still somewhat valuable?
Also, are we 100\% sure that this section is right?
A planet with larger surface field achieves an equivalent field to a planet with a smaller surface field at a larger radius, and captures a larger wind flux.
Should we think more about the spectrum?}

According to Eq.~(\ref{eq:Pkinp}), the total emitted power scales with $B^{2/3}$.
In detail, it is not obvious how the power spectral density scales with a planet's surface magnetic field, because the emission process involves electrons passing through a wide range of magnetic field conditions (dipolar field strength drops off as $r^{-3}$, and only in an idealized model is a planet a perfect magnetic dipole).
Nevertheless, a reasonable ansatz is that the radio emission is spread over a range of wavelengths in an approximately scale-independent way.
In other words, the frequency bandwidth $\Delta \nu$ of Eq.~(\ref{eq:Fnu}) is proportional to $f_{\rm cyc}$, as described in Section~\ref{ss:model_intensity} and in \citet{griesmeier2007b}.
This means that the power spectral density is likely to scale as the total power divided by the cyclotron frequency corresponding to a characteristic field strength where the emission peaks.
Since $f_{\rm cyc}$ scales as $B$ (a higher power of $B$ than the scaling of total power), this implies that the power spectral density in the radio scales as $B^{-1/3}$.
Planets that are intrinsically brighter, owing to a greater magnetic field strength, probably have their emission spread out over a broader set of frequencies, and therefore have a lower flux density as measured in Janskys!
\memoDS{Is this pointed out elsewhere in the literature?}

\section{Dave's sketch}

\citep{spiegel2008}

\citep{lecavelier_et_al2013}

\citep{janhunen_et_al2003}

\citep{zarka1992, zarka1998}

\citep{farrell_et_al2004}

\citep{lazio+farrell2007}: likelihood function, tau Boo search

\citep{lecavelier_et_al2009}

\citep{spiegel2012}

\citep{nordhaus+spiegel2013}

\citep{jiang+jin1996}: 12cm radio of Jupiter during Shoemake-Levy-9

\citep{morin2012, morin_et_al2013}

\citep{christensen_et_al2009, christensen2010}

\citep{saar2001}: Inverse rossby number scaling of magnetic field: $B
\sim 60 {\rm~G} \times Ro^{-1.2}$.  Here, he takes $Ro = \tau_c/P_{\rm
  rot}$, where $\tau_c$ is the convective turnover time = ?
\citep{gilliland1986}.  Well, $F \sim \rho v_c^3 \sim \rho
(H/\tau_c)^3$, so $\tau_c \sim H (\rho / F)^{1/3} = H / (\sigma T_{\rm
  eff}^4 / \rho)^{1/3}$.  Alternatively, $\tau_{\rm conv} \sim (M R^2
/ L)^{1/3}$.

On dimensional grounds, the convective turnover time goes as
$\tau_{\rm conv} \sim (M R^2 / L)^{1/3}$.  According to
\citet{burrows_et_al2001}, radius $R$ and luminosity scale with time
as $R \sim R_J$, where $R_J$ is the radius of Jupiter, and
\begin{equation}
\frac{L}{10^{-9} L_\odot} \sim \left( \frac{t}{1 \rm~Gyr} \right)^{-1.3} \left( \frac{M}{1~M_J} \right)^{2.64} \, .
\label{eq:burrowsLum_2}
\end{equation}
So,
%\begin{eqnarray}
%\nonumber \tau_{\rm conv} & \sim & 3 {\rm~days} \times \left( \frac{M}{M_J} \right)^{1/3} \left( \frac{R}{R_J} \right)^{2/3} \left( \frac{L}{L_\odot} \right)^{-1/3} \\
% & = & 
%\end{eqnarray}
\begin{eqnarray}
\nonumber \tau_c & \sim & 3 {\rm~hrs} \times \frac{\left( \frac{H}{100 \rm~km} \right) \left( \frac{\rho}{10^{-5} \rm~g~cm^{-3}} \right)^{1/3}}{\left( \frac{L}{10^{-9} L_\odot} \right)^{1/3}} \\
 & = & 3 {\rm~hrs} \times \frac{\left( \frac{H}{100 \rm~km} \right) \left( \frac{\rho}{10^{-5} \rm~g~cm^{-3}} \right)^{1/3}}{\left( \frac{M}{1 ~M_J} \right)^{0.88} \left( \frac{t}{1 \rm~Gyr} \right)^{0.43}} \\
\end{eqnarray}
So the Rossby number $Ro = \tau_c/P_{\rm rot}$ is
\begin{eqnarray}
Ro & = & \frac{\tau_c}{P_{\rm rot}} \\
 & = & \left( \frac{P_{\rm rot}}{3 \rm~hrs} \right)^{-1} \times \frac{\left( \frac{H}{100 \rm~km} \right) \left( \frac{\rho}{10^{-5} \rm~g~cm^{-3}} \right)^{1/3}}{\left( \frac{M}{1 ~M_J} \right)^{0.88} \left( \frac{t}{1 \rm~Gyr} \right)^{0.43}}
\end{eqnarray}

\citep{hallinan_et_al2013}

\citep{desch+kaiser1984} radiometric Bode's law

Noting that
\begin{equation}
\dot{M}_* = 4\pi r^2 \rho[r] v \, ,
\label{eq:mdot}
\end{equation}
so $\rho[r] = \dot{M}_*/(4\pi r^2 v)$,
\begin{eqnarray}
\frac{\rho v^2}{2} = \frac{B^2}{8\pi} \sim \frac{B_0^2}{8\pi} \left( \frac{d}{d_0} \right)^{-3} \, ,
\end{eqnarray}
where $B_0$ is the field strength at a distance $d_0$ from the planet.
\begin{eqnarray}
\frac{\rho v^2}{2} & \sim & \frac{B_0^2}{8\pi} \left( \frac{d}{d_0} \right)^{-6} \\
\frac{\dot{M}_* v}{8\pi a^2} & = & \frac{B_0^2}{8\pi} \left( \frac{d}{d_0} \right)^{-6} \\
\frac{\dot{M}_* v}{r^2} & = & B_0^2 \left( \frac{d}{d_0} \right)^{-6} \\
d^6 & = & \frac{d_0^6 B_0^2 r^2}{\dot{M}_* v} \\
d_A & = & d_0 \left( \frac{B_0^2 r^2}{\dot{M}_* v} \right)^{1/6} \\
  & \sim & 4 R_J \left( \frac{d_0}{R_J}\right) \left\{ \frac{\left( \frac{B}{10 \rm~G} \right)^2 \left( \frac{r}{5 \rm~AU} \right)^2}{\left( \frac{\dot{M}_*}{10^{-6} M_\odot/\rm yr} \right) \left( \frac{v}{20 \rm~km/s} \right)} \right\}^{1/6}
\label{eq:Chapman-Ferraro}
\end{eqnarray}
In the above, $d_A$ is the distance from the planet to the Alfven
point, where the magnetic energy density $u_B = B^2 / 8\pi$ equals the
kinetic energy density in the stellar wind $u_w = \rho v^2/2$, and
$B_0$ is the magnetic field strength at distance $d_0$.

The escape speed is
\begin{equation}
v_{\rm esc}^* = \sqrt{\frac{2 G M_*}{R_*}} \, ,
\end{equation}
so if the stellar wind speed is a factor $C$ times the escape speed, then ...

The power incident on the planet within $d_A$ is
\begin{eqnarray}
P_{\rm inc} & = & \frac{\rho v^3}{2} \times \pi d_A^2 \\
\frac{\rho v^2}{2} & \sim & \frac{B_0^2}{8\pi} \left( \frac{d}{d_0} \right)^{-6} \\
d_A^6 & = & d_0^6 \frac{B_0^2}{4\pi \rho v^2} \\
d_A^2 & = & d_0^2 \left( \frac{B_0^2}{4\pi \rho v^2}\right)^{1/3} \\
\pi d_A^2 \frac{\rho v^3}{2} & = & \frac{d_0^2}{2} \left( \frac{\pi^2 \rho^2 v^7 B_0^2}{4} \right)^{1/3} \\
P_{\rm inc} & = & d_0^2 \left( \frac{\pi^2 \rho^2 v^7 B_0^2}{32} \right)^{1/3} 
%\frac{\dot{M}_* v}{8\pi a^2} & = & \frac{B_0^2}{8\pi} \left( \frac{d}{d_0} \right)^{-6} \\
%\frac{\dot{M}_* v}{r^2} & = & B_0^2 \left( \frac{d}{d_0} \right)^{-6} \\
%d^6 & = & \frac{d_0^6 B_0^2 r^2}{\dot{M}_* v} \\
%d_A & = & d_0 \left( \frac{B_0^2 r^2}{\dot{M}_* v} \right)^{1/6} \\
%  & \sim & 4 d_0 \left\{ \frac{\left( \frac{B}{10 \rm~G} \right)^2 \left( \frac{r}{5 \rm~AU} \right)^2}{\left( \frac{\dot{M}_*}{10^{-6} M_\odot/\rm yr} \right) \left( \frac{v}{20 \rm~km/s} \right)} \right\}^{1/6}
\end{eqnarray}
Note that $\rho v = \dot{M}_*/4\pi r^2$.  Therefore,
\begin{eqnarray}
P_{\rm inc} & = & d_0^2 \left( \frac{\pi^2 \rho^2 v^7 B_0^2}{32} \right)^{1/3} \\
 & = & d_0^2 \left( \frac{\dot{M}_*^2 v^5 B_0^2}{512 r^4} \right)^{1/3} \\
 & = & \frac{d_0^2}{8} \left( \frac{\dot{M}_*^2 v^5 B_0^2}{r^4} \right)^{1/3} \\
\nonumber  & \sim & 2 \times 10^{18} {\rm~W} \left( \frac{d_0}{R_J} \right)^2 \left( \frac{\dot{M}_*}{10^{-5} M_\odot / {\rm yr}} \right)^2 \\
 & & \times \left( \frac{v}{20 \rm~km/s} \right)^5 \left( \frac{B_0}{10 \rm~G} \right)^2 \left( \frac{r}{5 \rm~AU} \right)^{-4}
\end{eqnarray}

The power incident on the planet is

%\citep{lunine_et_al1989} -- error?

% http://kiss.caltech.edu/workshops/magnetic2013/presentations/winterhalter.pdf

%ftp://ftp.iwf.oeaw.ac.at/pub/Scherf/PRE-CD%20V2/pre6/nigl.pdf

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Assumptions for Planetary Magnetic Field}
%\label{ss:magneticfield}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\memoYF{It may be better to consider Cristensen's scaling, taking account of the age. But I have not followed the theory yet.}

Theoretically, the magnetic moment of gaseous planets are expressed with the following scaling relationship \citep{griesmeier2004}:
%%%%%%%%%% 
\begin{equation}
\mathcal{M} \propto  \omega ^{\alpha } \rho_c ^{\beta } r_c^{\gamma } \sigma ^{\delta }
\end{equation}
%%%%%%%%%%
where $\omega $ is the spin angular velocity, $\rho _c$, $r_c$ and $\sigma $ are the density, the radius, and the conductivity of the ``dynamo region'' where the density is high enough for hydrogen to be metallic, respectively. 
The scaling indexes are estimated to be $\alpha \sim 1/2-1$, $\beta \sim 1/2$, $\gamma \sim 3-4$, and $\sigma \sim -1/2-0$. In this paper, we assume $\alpha =1$, $\beta =1/2$, and $r_c = 7/2$ \citep{sano1993}. \memoYF{validity?}

Unlike usual hot jupiters, RGHJs are not subject to tidal lock, as the gravitational effects of their host star does not change even if the star evolves into red giants. Without no physical insights of the typical spin rate, we simply assume that of Jupiter: $\omega = 9.925$ [hr]. 
\memoYF{Would their spin rate become higher or lower??}

In order to evaluate $\rho _c $ and $r_c$, we need a model of internal structure of gaseous planets. 
First, we assume that the planetary radius is constant at $R_p = R_{p,J}$, as the numerical calculations show that the radii of gaseous planets over the range of $0.1 R_{p, J} < M_p < 10M_{p, J}$ (with core mass less than 10\%) are converged to $0.8 R_{p, J} < R_p < 1.2R_{p, J}$ in 1 Gyr \citep{fortney2007}. 
For the density profile, we assume a polytrope gas sphere with index $n=1$, which results in:
%%%%%%%%%% 
\begin{equation}
\rho [r] = \left( \frac{\pi M_p}{4 R_p^3} \right) \frac{\sin \left[ \pi \frac{r}{R_p} \right]}{\left( \pi \frac{r}{R_p} \right)}. \label{eq:rho_r_2}
\end{equation}
%%%%%%%%%%
We determine the core radius $r_c$ by assuming that the hydrogen becomes metallic when $\rho (r)$ exceeds the critical density $\rho_c=700\,\mbox{kg/m}^3$. The density of the metallic core, $\rho _c$ is obtained by averaging the density in the core. 
We assume that the conductivity $\sigma $ is the same as Jupiter \memoYF{?}.

\end{document}